Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-17 analysis/usage.Rmd
usage

```{r read, message=TRUE, include=F}
# This test should allow to re-run the whole script without reading the
# data again from disk. It assumes the data doesn't change.
if (exists(eval_calls_path) == FALSE) {
  read_fst(eval_calls_path) %>% as_tibble() -> E_raw
  read_fst(corpus_path) %>% as_tibble() -> C_raw
  read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
}
```

@@

E %>% filter(is.na(arg_size) == FALSE) %>%
  filter(arg_size > 100) %>%
  ggplot(aes(x=arg_size)) + geom_histogram(binwidth=1) 
```

@@

# Amount of computations via eval

```{r amount_computations, include=F}
nb_events_in_eval <- E %>% summarise(nb_events = sum(opcodes))

max_events_in_eval <- E %>% summarize(maxi = max(opcodes))

less_100_events <- E %>% filter(opcodes <= 100)
nb_less_100_events <- count(less_100_events, wt = duplicates)$n

r("nb events eval", nb_events_in_eval$nb_events)
r("maxi events eval", max_events_in_eval$maxi)
r("small events eval", ratio(nb_less_100_events, nb_calls))
```

```{r nb-events-violin-plot,include=F}
E %>%
  filter(opcodes > 0) %>%
  ggplot(aes(x = "", y = opcodes, fill = ""), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "Number of events in eval per package") +
  theme(legend.position = "none")

ggsave(path(PLOT_DIR, "nb_events_per_packages.pdf"))
```

@@

```{r loops}
eval_calls_exercised <- E %>%
  group_by(file) %>%
  count(src_ref, wt = duplicates)

eval_calls_exercised %>% ggplot(aes(n)) +
  geom_histogram() +
  scale_x_log10(labels = scales::comma) +
  labs(x = "Number of eval calls per srcref per run")
```

```{r}
eval_call_triggered <- E %>%
  group_by(src_ref) %>%
  summarize(n_runs = n_distinct(file), n = sum(duplicates), average = n / n_runs) %>%
  arrange(desc(average))

eval_call_triggered %>% 
  filter(average < 50) %>% 
  ggplot(aes(x = average)) +
  geom_histogram(binwidth =  1)

@@

bins <- cut(eval_call_triggered$average, breaks = c(0, 50, 100, 250,500, 1000, 1500,  2000, 3000), include.lowest = TRUE)
sum <- summary(bins)

@@

```{r}
duration <- difftime(Sys.time(), now)
```

>>>>

```{r read, message=TRUE, include=F}
# This test should allow to re-run the whole script without reading the
# data again from disk. It assumes the data doesn't change.
if (exists("E") == FALSE) {
  read_fst(eval_calls_path) %>% as_tibble() -> E_raw
  read_fst(corpus_path) %>% as_tibble() -> C_raw
  read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
}
```

@@

sizes_prop %>%
  ggplot(aes(x=arg_size)) + geom_histogram()

```

@@

# Amount of computation

How much work is involved in each eval as measured by the number of ??? 
(possible: opcodes, AST nodes).

```{r amount_computations, include=F}
tot_events <- E %>% summarise(n=sum(opcodes)) %>% .$n
max_events <- E %>% summarize(max = max(opcodes)) %>% .$n

less_50_events <- E %>% filter(opcodes <= 50) %>%
  count(wt = duplicates) %>%
  .$n

overview_table(
  r("Events", tot_events),
  r("Maxevents", max_events),
  r("Smalleventspct", ratio(less_50_events, nb_calls))
) -> TBL

E %>%
  filter(opcodes > 50) %>%
  ggplot(aes(x = "", y = opcodes, fill = ""), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "", x="") +
  theme(legend.position = "none")

ggsave(path(PLOT_DIR, "events_per_pack_large.pdf"))

E %>%
  filter(opcodes > 0) %>%
  filter(opcodes < 50) %>%
  ggplot(aes(x = "", y = opcodes, fill = ""), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7) +
  labs(y = "Events per package", x="") +
  theme(legend.position = "none")

ggsave(path(PLOT_DIR, "events_per_pack_small.pdf"))
```

@@

```{r loops, include=FALSE}
# Which call sites where exercised and how many times, per run.
exercised <- E %>%
  group_by(file) %>%               # group by runs
  count(src_ref, wt = duplicates)  # weigh by duplicates

exercised %>% ggplot(aes(n)) +
  geom_histogram() +
  scale_x_log10(labels = scales::comma) +
  labs(x = "Calls per site per run")

# For each site, how many runs and within each runs how often do we trigger.
triggered <- E %>%
  group_by(src_ref) %>%
  summarize(n_runs = n_distinct(file), 
            n = sum(duplicates), 
            average = n / n_runs) %>%
  arrange(desc(average))

# Plot the low trigger counts...
triggered %>% 
  filter(average < 50) %>% 
  ggplot(aes(x = average)) +
  geom_histogram(binwidth =  1)

@@

bins <- cut(triggered$average, breaks = c(0, 50, 100, 250,500, 1000, 1500,  2000, 3000), include.lowest = TRUE)
sum <- summary(bins)

@@

```{r include=FALSE}
duration <- difftime(Sys.time(), now)
```

### 2020-12-17 paper/main.tex
usage

. We turn off the
bytecode compiler for this study. The bytecode compiler can also call
\eval. We do not get source locations for \UndefinedEvalsRnd \eval
calls. In these cases \eval is either passed as an argument to a
higher-order functions or is defined in a function returned by a
higher-order function and the R parser does not retain location
information for \eval. However, this is a meager \PercentUndefinedEval
of all \eval calls and is unlikely to affect our analysis. We ignore
calls to the native \eval function exposed by R. We also ignore the
\c{rlang::tidy_eval} function which uses native \eval internally
because \c{rlang} is used to implement a DSL for data analysis in R.
It introduces a new first-class promise object called \c{quosure} for
which it implements special evaluation support in \c{tidy_eval}.

@@

\begin{figure}[h] \centering
    \includegraphics[width=\textwidth]{calls_per_run_per_call_site}
    \caption{Invocations of \eval per call sites normalized by the
      number of runs, for less than 50 normalized invocations. Binsize
      is 1} \label{fig:callsitesnormalized}
\end{figure}

@@

\mypara{Amount of computation via eval.} We compare the number of
events intercepted during execution of \eval across the whole corpus.
For this, we use the number of interpreter steps as a proxy for
events. We observe \AllEventCountRnd{} \NOTE{Check that number.\quad}
events all told, and \NbEventsEvalRnd events occur inside eval
(\EvalEventAllPerc of all events). Figure~\ref{fig:eventsPerPackage}
show a violin plot of events.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{nb_events_per_packages}
	\caption{Violin plot of the number of events per packages (log scale). } \label{fig:eventsPerPackage}
\end{figure}

We observe a very wide spread in the number of events generated by
\eval calls. The largest number of events generated by an \eval call
is \MaxiEventsEvalRnd. However, a majority (\SmallEventsEval) of \eval
calls performs up to 100 events suggesting that most expressions
passed to \eval are small. It is interesting to note that a higher
proportion of Core R \eval calls perform side effects in the
\EventsMinRangeB--\EventsMaxRangeB range. Only \EventsCoreEvalCountC
Core R \eval calls and \EventsPackageEvalCountC package \eval call
generate \EventsMinRangeC to \EventsMaxRangeC events. These \eval
calls originate from statistical modeling packages such as \mlogit,
\mboost, \metafor, \lavaan, \mclust and \gamlss.

>>>>

We turn off the bytecode compiler for this study. The bytecode
compiler can also call \eval. We do not get source locations for
\UndefinedEvalsRnd \eval calls. In these cases \eval is either passed
as an argument to a higher-order functions or is defined in a function
returned by a higher-order function and the R parser does not retain
location information for \eval. However, this is a meager
\PercentUndefinedEval of all \eval calls and is unlikely to affect our
analysis. We ignore calls to the native \eval function exposed by R.
We also ignore the \c{rlang::tidy_eval} function which uses native
\eval internally because \c{rlang} is used to implement a DSL for data
analysis in R. It introduces a new first-class promise object called
\c{quosure} for which it implements special evaluation support in
\c{tidy_eval}.

@@

\begin{figure}[h] \centering
    \includegraphics[width=.5\textwidth]{calls_per_run_per_call_site}
    \caption{Invocations of \eval per call sites normalized by the
      number of runs, for less than 50 normalized invocations. Binsize
      is 1} \label{fig:callsitesnormalized}
\end{figure}

@@

\mypara{Amount of computation via eval.} We compare the number of
events intercepted during execution of \eval across the whole corpus.
For this, we use the number of interpreter steps as a proxy for
events. We observe \AllEventCountRnd \NOTE{Check that number.}
events all told, and \EventsRnd events occur inside eval
(\EvalEventAllPerc of all events). Figure~\ref{fig:eventsPerPackage}
show a violin plot of events.

\begin{figure}[h]
	\centering
	\includegraphics[width=.4\textwidth]{events_per_pack_small}
	\includegraphics[width=.4\textwidth]{events_per_pack_large}
	\caption{Events per packages} \label{fig:eventsPerPackage}
\end{figure}

We observe a very wide spread in the number of events generated by
\eval calls. The largest number of events generated by an \eval call
is \MaxeventsRnd. However, a majority (\Smalleventspct) of \eval calls
performs up to 50 events suggesting that most expressions passed to
\eval are small. It is interesting to note that a higher proportion of
Core R \eval calls perform side effects in the
\EventsMinRangeB--\EventsMaxRangeB range. Only \EventsCoreEvalCountC
Core R \eval calls and \EventsPackageEvalCountC package \eval call
generate \EventsMinRangeC to \EventsMaxRangeC events. These \eval
calls originate from statistical modeling packages such as \mlogit,
\mboost, \metafor, \lavaan, \mclust and \gamlss.
